---
title: "HuYuDataInsight LLC"
author: "Zhaowei Cai"
date: "2024-05-09"
output:
  pdf_document: default
  html_document: default
---

**(a)**
```{r}
library(urca)
library(vars)
library(quantmod)
library(forecast)
library(tseries)
library(fGarch)
library(stargazer)


data(denmark)
df.lev = denmark[,c('LRM','LRY','IBO','IDE','LPY')]
m.lev  = as.matrix(df.lev)
nr_lev = nrow(df.lev)
 
# quarterly centered dummy variables
dum_season = data.frame(yyyymm = denmark$ENTRY)
substr.q = as.numeric(substring(denmark$ENTRY, 6,7))
dum_season$Q2 = (substr.q==2)-1/4
dum_season$Q3 = (substr.q==3)-1/4
dum_season$Q4 = (substr.q==4)-1/4
dum_season = dum_season[,-1]

# visualization
str.main = c(
    'LRM=ln(real money M2)', 'LRY=ln(real income)',
    'IBO=bond rate', 'IDE=bank deposit rate', 'LPY=ln(price deflator)')
par(mfrow=c(3,2), mar=c(5,3,3,3))
for(i in 1:5) {
    matplot(m.lev[,i], axes=FALSE,
        type=c('l'), col = c('blue'), 
        main = str.main[i]) # plot the columns of one matrix m.lev
    axis(2) # y axis
    axis(1, at=seq_along(1:nrow(df.lev)),
         labels=denmark$ENTRY, las=2)
}
```

**(b)**
```{r}
##Check for the trend (the Augmented Dickey–Fuller (ADF) test)
r1 = df.lev$LRM
acf(r1)
pacf(r1)
adf.test(r1)
summary(ur.df(r1, type='trend', lags=20, selectlags="BIC"))
# drift+difference
diffr1 = na.omit(diff(r1))
fit1 = auto.arima(diffr1)
summary(fit1)
tsdiag(fit1)

##Check for the trend (the Augmented Dickey–Fuller (ADF) test)
r2 = df.lev$LRY
acf(r2)
pacf(r2)
adf.test(r2)
summary(ur.df(r2, type='trend', lags=20, selectlags="BIC"))
# difference
diffr2 = na.omit(diff(r2))
fit2 = auto.arima(diffr2)
summary(fit2)
tsdiag(fit2)

##Check for the trend (the Augmented Dickey–Fuller (ADF) test)
r3 = df.lev$IBO
acf(r3)
pacf(r3)
adf.test(r3)
summary(ur.df(r3, type='trend', lags=20, selectlags="BIC"))
# drift+tt+difference
diffr3 = na.omit(diff(r3))
summary(ur.df(diffr3, type='trend', lags=20, selectlags="BIC"))
fit3 = auto.arima(diffr3)
summary(fit3)
tsdiag(fit3)

##Check for the trend (the Augmented Dickey–Fuller (ADF) test)
r4 = df.lev$IDE
acf(r4)
pacf(r4)
adf.test(r4)
summary(ur.df(r4, type='trend', lags=20, selectlags="BIC"))
# drift+tt+difference
diffr4 = na.omit(diff(r4))
summary(ur.df(diffr4, type='trend', lags=20, selectlags="BIC"))
fit4 = auto.arima(diffr4)
summary(fit4)
tsdiag(fit4)

##Check for the trend (the Augmented Dickey–Fuller (ADF) test)
r5 = df.lev$LPY
acf(r5)
pacf(r5)
adf.test(r5)
summary(ur.df(r5, type='trend', lags=20, selectlags="BIC"))
diffr5 = na.omit(diff(r5))
fit5 = auto.arima(diffr5)
summary(fit5)
tsdiag(fit5)
```

**(c)**
```{r}
res=residuals(fit1)
shapiro.test(res)
par(mfrow=c(2,1))
hist(res)
lines(density(res))
qqnorm(res)          
qqline(res)

res=residuals(fit2)
shapiro.test(res)
par(mfrow=c(2,1))
hist(res)
lines(density(res))
qqnorm(res)          
qqline(res)

res=residuals(fit3)
shapiro.test(res)
par(mfrow=c(2,1))
hist(res)
lines(density(res))
qqnorm(res)          
qqline(res)

res=residuals(fit4)
shapiro.test(res)
par(mfrow=c(2,1))
hist(res)
lines(density(res))
qqnorm(res)          
qqline(res)

res=residuals(fit5)
shapiro.test(res)
par(mfrow=c(2,1))
hist(res)
lines(density(res))
qqnorm(res)          
qqline(res)
```

**(d)**
```{r}
# LRM
p1 <- forecast(fit1, h=3, level=0.95)
last1 <- r1[length(r1)]
(LRM1 <- last1+cumsum(p1$mean))
# 95% CI:
last1+cumsum(p1$lower) # lb
last1+cumsum(p1$upper) # ub
# LRY
p2 <- forecast(fit2, h=3, level=0.95)
last2 <- r2[length(r2)]
(LRY1 <- last2+cumsum(p2$mean))
# 95% CI:
last2+cumsum(p2$lower) # lb
last2+cumsum(p2$upper) # ub
# IBO
p3 <- forecast(fit3, h=3, level=0.95)
last3 <- r3[length(r3)]
(IBO1 <- last3+cumsum(p3$mean))
# 95% CI:
last3+cumsum(p3$lower) # lb
last3+cumsum(p3$upper) # ub
# IDE
p4 <- forecast(fit4, h=3, level=0.95)
last4 <- r4[length(r4)]
(IDE1 <- last4+cumsum(p4$mean))
# 95% CI:
last4+cumsum(p4$lower) # lb
last4+cumsum(p4$upper) # ub
# LPY
p5 <- forecast(fit5, h=3, level=0.95)
last5 <- r5[length(r5)]
(LPY1 <- last5+cumsum(p5$mean))
# 95% CI:
last5+cumsum(p5$lower) # lb
last5+cumsum(p5$upper) # ub
```

2. Please (a) build the best VAR models using all five variables together. (b) Please be sure to check the integrated order, and make considerations for trend and seasonality. (c) Please check the residuals to ensure your model is a good fit. (d). Please use your best VAR model to make forecast for the next three quarters.

```{r}
df.diff = diff(as.matrix(df.lev), lag = 1)
colnames(df.diff) = c('dLRM', 'dLRY', 'dIBO', 'dIDE', 'dLPY')
m.diff = as.matrix(df.diff)
# lag length
VARselect(df.diff, lag.max = 4, type = 'const', season = 4)
# estimation
vare_diff = VAR(df.diff, p = 1, type = 'const', season = 4)
summary(vare_diff)
# residuals test
serial.test(vare_diff)
# forecast of differenced data
varf_diff = predict(vare_diff, n.ahead = 3, ci = 0.95)
# predictions & 95%CI
(LRM2 <- last1+cumsum(varf_diff$fcst$dLRM[,1]))
last1+cumsum(varf_diff$fcst$dLRM[,2]) # lb
last1+cumsum(varf_diff$fcst$dLRM[,3]) # ub
(LRY2 <- last2+cumsum(varf_diff$fcst$dLRY[,1]))
last2+cumsum(varf_diff$fcst$dLRY[,2]) # lb
last2+cumsum(varf_diff$fcst$dLRY[,3]) # ub
(IBO2 <- last3+cumsum(varf_diff$fcst$dIBO[,1]))
last3+cumsum(varf_diff$fcst$dIBO[,2]) # lb
last3+cumsum(varf_diff$fcst$dIBO[,3]) # ub
(IDE2 <- last4+cumsum(varf_diff$fcst$dIDE[,1]))
last4+cumsum(varf_diff$fcst$dIDE[,2]) # lb
last4+cumsum(varf_diff$fcst$dIDE[,3]) # ub
(LPY2 <- last5+cumsum(varf_diff$fcst$dLPY[,1]))
last5+cumsum(varf_diff$fcst$dLPY[,2]) # lb
last5+cumsum(varf_diff$fcst$dLPY[,3]) # ub
```
3.
we have already got forecast for the next 3 quarters based on two methods. If we know the true values of the following 3 quarters, we can compute the MSE to compare which one is better.

```